# How I read? - Introduction to Scoopinion reading behavior data with R
### Juuso Parkkinen - @ouzor
### 15.6.2012

```{r setup, include=FALSE}
# upload images to imgur automatically 
opts_knit$set(upload.fun = imgur_upload)

# Load required libraries
library(ggplot2)
theme_set(theme_grey(16))

```

---

### Reading the data

Read the JSON data file

```{r read_data, warning=FALSE}
# Use package 'rjson' to read the data
library(rjson)
scoop.raw <- fromJSON(file="myscoopiniondata.json")
```

See what metadata is given

```{r meta}
# Check meta data
scoop.raw$meta
```
---

### Preprocessing
The interesting part is the 'reads' data. Process it into analysable format:

```{r preprocess, warning=FALSE}
common.items <- c("id", "words_read", "created_at", "article.id", "article.title", "article.url", "article.word_count", "article.average_time", "article.language", "article.site.name", "article.site.id", "article.site.url")
temp.mat <- sapply(scoop.raw$reads, function(x) {res=unlist(x); return(res[names(res) %in% common.items])})
# Produce a data frame
scoop.df <- data.frame(t(temp.mat))
# Transform some columns to numeric form
scoop.df[c("words_read", "article.word_count", "article.average_time")] <- sapply(scoop.df[c("words_read", "article.word_count", "article.average_time")], function(x) as.numeric(as.vector(x)))
# Remove reads with missing data
scoop.df <- na.omit(scoop.df)

```

R has nice tools for handling Dates, with easy addition of weekday and month information:

```{r date}
# Add date in a proper format
scoop.df$Date <- as.Date(sapply(strsplit(as.vector(scoop.df$created_at), split="T"), function(x) x[1]))
# Add weekdays
scoop.df$WeekDay <- factor(weekdays(scoop.df$Date), levels=c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))
# Add months
scoop.df$Month <- factor(months(scoop.df$Date), levels=month.name)
# Add years
scoop.df$Year <- format(scoop.df$Date, format = "%Y")
# Add combined year-month information
scoop.df$Year.Month <- factor(paste(scoop.df$Year, scoop.df$Month, sep=" - "))
temp.ym <- paste(rep(c("2012", "2013"), each=length(month.name)), month.name, sep=" - ")
scoop.df$Year.Month <- factor(scoop.df$Year.Month, levels=temp.ym)
# Cut away 2013 - April (only one read)
scoop.df <- droplevels(subset(scoop.df, Year.Month != "2013 - April"))
```

---
### Basic stuff

```{r basic}
# Most commonly read sites
head(sort(table(scoop.df$article.site.name), decreasing=T), 10)

# Language distribution
table(scoop.df$article.language)

# Remove rare languages
scoop.df <- droplevels(subset(scoop.df, article.language %in% c("en", "fi")))

# Top referring sites (currently not included in the anlysis because they are missing from some reads)
# head(sort(table(scoop.df$referrer), decreasing=T), 5)
```


---
### Plot word counts against average reading times

```{r word_vs_time, message=FALSE, fig.width=8, fig.heigth=8}
# Scatter plot of word count vs. average time, colour based on languages
ggplot(scoop.df, aes(x=article.word_count, y=article.average_time, colour=article.language)) + geom_jitter()
```

---
### Plot reading behaviour over time, split by years

```{r time, fig.width=10, fig.height=6}
# Histogram of daily reading counts
ggplot(scoop.df, aes(x=Date, fill=article.language)) + geom_histogram(position="stack", binwidth=1)+ facet_wrap(~ Year, ncol=1, scales="free_x")
```

---
### Plot reading counts for different weekdays


```{r weekday, fig.width=12, fig.height=6}
# Histogram of weekday reading counts
ggplot(scoop.df, aes(x=WeekDay, fill=article.language)) + geom_histogram(position="stack", binwidth=1) + facet_wrap(~ Year.Month, ncol=5) + theme(axis.text.x=element_text(angle=45, vjust=0.8))
```

