# How I read? - Introduction to Scoopinion reading behavior data with R
### Juuso Parkkinen - @ouzor
### 16.6.2012

```{r setup, include=FALSE}
# upload images to imgur automatically 
opts_knit$set(upload.fun = imgur_upload)

# Load required libraries
library(plyr)
library(ggplot2)
theme_set(theme_grey(16))

```

---

### Read and preprocess data

Read the JSON data file and preprocess reads data (see details in the [source code](scoopinion_functions.R))

```{r read_data, warning=FALSE, message=FALSE}
source("scoopinion_functions.R")
reads.df <- PreprocessScoopinionData(filename="my_scoopinion_data.json")
```


---
### Filter data

Read code comments for details

```{r filter, fig.width=6, fig.height=4}
# Remove rare languages
table(reads.df$article.language)
reads.df <- droplevels(subset(reads.df, article.language %in% c("en", "fi")))

# Remove rare months
table(reads.df$Year.Month)
reads.df <- droplevels(subset(reads.df, Year.Month != "2013-04"))

# Study reads with repeating 'article.title'
repeats <- table(reads.df$article.title)
repeats <- repeats[repeats > 2]
print(repeats)
# Based on this, remove comics and Yle Areena
reads.to.remove <- which(reads.df$article.title %in% c("Anonyymit eläimet", "Fok_it", "Fok_it — Nyt", "Sarjakuvat", "Yle Areena"))
message("Removed ", length(reads.to.remove), " reads")
reads.df <- droplevels(reads.df[-reads.to.remove, ])

# Remove very long articles
ggplot(reads.df, aes(x=article.word_count)) + geom_histogram(binwidth=500)
reads.df <- droplevels(subset(reads.df, article.word_count < 5000))
```

---
### Top sites

List top read article sites and to referrers (the site from which the read article was clicked)

```{r top_sites}
# Most commonly read sites
head(sort(table(reads.df$article.site.name), decreasing=T), 10)

# Top referring sites
head(sort(table(reads.df$referrer), decreasing=T), 10)
```

---
### Study reading time vs. word count

Scatter plot of average reading time vs. word count, with estimated means. 

```{r word_vs_time1, warning=FALSE, message=FALSE, fig.width=10, fig.heigth=8}
ggplot(reads.df, aes(x=article.word_count, y=article.average_time, colour=article.language)) + geom_point(position=position_jitter(width=0, height=10)) + geom_smooth(data=subset(reads.df, article.average_time>0))
```

My personal reading time vs. words.read

```{r word_vs_time2, warning=FALSE, message=FALSE, fig.width=10, fig.heigth=8}
ggplot(reads.df, aes(x=words_read, y=total_time, colour=article.language)) + geom_point()
```

My personal reading time vs. article word count

```{r word_vs_time3, warning=FALSE, message=FALSE, fig.width=10, fig.heigth=8}
ggplot(reads.df, aes(x=article.word_count, y=total_time, colour=article.language)) + geom_point()
```

*COMMENT: There appears to be a threshold for too high reading speed (a bit higher than 4 words / second)*
*COMMENT: There are also a lot of reads with zero words read, probably some measuring errors. But why some articles have 'average_time' zero?*

Personal words.read vs. article word count
```{r word_vs_time4, warning=FALSE, message=FALSE, fig.width=10, fig.heigth=8}
ggplot(reads.df, aes(x=article.word_count, y=words_read, colour=article.language)) + geom_abline(slope=1, linetype="dashed") + geom_jitter()
```

Compare my reading speed (total_time) to average

```{r word_vs_time5, warning=FALSE, message=FALSE, fig.width=10, fig.heigth=8}
ggplot(reads.df, aes(x=article.average_time, y=total_time, colour=article.language)) + geom_abline(slope=1, linetype="dashed") + geom_jitter()
```


---
### Plot reading behaviour over time

Histogram of daily reads count

```{r time1, fig.width=10, fig.height=6}
# Histogram of daily reading counts
ggplot(reads.df, aes(x=Date, fill=article.language)) + geom_histogram(position="stack", binwidth=1)+ facet_wrap(~ Year, ncol=1, scales="free_x")
```

Histogram of reading counts by weekdays, split by months

```{r time2, fig.width=10, fig.height=6}
# Histogram of daily reading counts
ggplot(reads.df, aes(x=WeekDay, fill=article.language)) + geom_histogram(position="stack", binwidth=1) + facet_wrap(~ Year.Month, ncol=5) + theme(axis.text.x=element_text(angle=45, vjust=0.8))
```

Study monthly averages for 10 most read sites. Add vertical line for HS paywall, introduced 20.11.2013.

```{r time3, fig.width=10, fig.height=6}
top10.sites <- names(head(sort(table(reads.df$article.site.name), decreasing=T), 10))
top10.df <- droplevels(subset(reads.df, article.site.name %in% top10.sites))
# Compute monthly averages
top10.montly.df <- plyr::ddply(top10.df, c("article.site.name", "Year.Month"), summarise, Average_reads=length(id))

# Position of HS paywall
paywall.pos <- which(unique(top10.montly.df$Year.Month)=="2012-11")
ggplot(top10.montly.df, aes(x=Year.Month, y=Average_reads, colour=article.site.name)) + geom_path(aes(group=article.site.name)) + geom_vline(xintercept=paywall.pos, linetype="dashed") + annotate("text", x=paywall.pos+0.1, y=40, label="HS paywall introduced 20.11.2012", hjust=0)
```


